An Analysis of daily shelter and overnight service usage in Toronto focusing on shelter occupancy and capacity

### 2024-09-27 paper/paper.qmd
Final Revision
Correct data style
Applying styler for all codes
Correct code styles for the paper

## Data Tools
The data was simulated, extracted, and cleaned by R[@citeR] from the opendatatoronto package[@citeopendatatoronto]. Tidyverse[@citetidyverse], here[@citehere], ggplot2[@citeggplot2], dplyr[@citedplyr], lubridate[@citelubridate], janitor[@citejanitor] and knitr[@citeknitr] are all applied in the analysis.

@@

#  Create a subset includes key variables
subset <- data[, c("year", "overnight_service_type", "service_user_count", "occupancy_rate_beds", 
                   "occupancy_rate_rooms")]

@@

# Present the sample as a table
kable(sample, col.names = c("year", "overnight service type", "service user count", 
                            "occupancy rate beds", "occupancy rate rooms"), 
      align = "ccc", 
      caption = "Table 1: Sample of the key variables in Occupancy & Capacity Data")
```

@@

#  Create a subset includes key variables
summary_stats <- data.frame(Variable = c("Service User Count", "Occupancy Rate of Beds", 
                                         "Occupancy Rate of Rooms"), 
                            Mean = c(mean_user_cnt, mean_rate_bed, mean_rate_room), 
                            SD = c(sd_user_cnt, sd_rate_bed, sd_rate_room))

# Present the sample as a table
kable(summary_stats, col.names = c("Variable", "Mean", "Standard Deviation"), 
      align = "ccc", 
      caption = "Table 2: Mean and Standard Deviation of Key Variables")
```

@@

#  Create a subset includes key variables
summary_stats <- data.frame(Year = c("2021", "2022", "2023", "2024"), 
                            Service_User_Counts_Sum = c(sum_cnt_2021, sum_cnt_2022, 
                                                       sum_cnt_2023, sum_cnt_2024), 
                            Actual_Bed_Capacity_Sum = c(sum_bed_2021, sum_bed_2022, 
                                                        sum_bed_2023, sum_bed_2024), 
                            Actual_Room_Capacity_Sum = c(sum_room_2021, sum_room_2022, 
                                                         sum_room_2023, sum_room_2024)
                            )

# Present the sample as a table
kable(summary_stats, col.names = c("Year", "Total Service User Counts", 
                                   "Total Actual Bed Capacity", "Total Actual Room Capacity"), 
      align = "ccc", 
      caption = "Table 3: Total Service User Count and Actual Capacity between years")
```

@@

data |> ggplot(aes(x = overnight_service_type, y = service_user_count)) +  
  geom_col(fill = "orange") + 
  theme_minimal() + 
  labs(x = "overnight service types", y = "service user counts") + 
  coord_flip() + 
  scale_y_continuous(labels = scales::comma)

```

@@

# Reshape the data to long format for plotting with side-by-side bars
df_long <- tidyr::pivot_longer(avg_rate, cols = c("avg_rate_bed", "avg_rate_room"),
                               names_to = "type",
                               values_to = "average")

# Plot using ggplot2
ggplot(df_long, aes(x = factor(year), y = average, fill = type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  labs(x = "Year", y = "Average Occupancy Rate", fill = "Type") +
  theme_minimal() + 
  geom_text(aes(label = round(average, 1)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 3.5)

```

@@

# Reshape the data to long format for plotting with side-by-side bars
df_long <- tidyr::pivot_longer(total_cap, cols = c("total_cap_bed", "total_cap_room"),
                               names_to = "type",
                               values_to = "total")

# Plot using ggplot2
ggplot(df_long, aes(x = factor(year), y = total, fill = type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  labs(x = "Year", y = "Total Actual Capacity", fill = "Type") +
  theme_minimal()

```

>>>>

## Data Tools
The data was simulated, extracted, and cleaned by R[@citeR] from the opendatatoronto package[@citeopendatatoronto]. Tidyverse[@citetidyverse], here[@citehere], ggplot2[@citeggplot2], dplyr[@citedplyr], lubridate[@citelubridate], janitor[@citejanitor] and knitr[@citeknitr] are all applied in the analysis. Styler[@citestyler] is used to correct code styles.

@@

#  Create a subset includes key variables
subset <- data[, c(
  "year", "overnight_service_type", "service_user_count", "occupancy_rate_beds",
  "occupancy_rate_rooms"
)]

@@

# Present the sample as a table
kable(sample,
  col.names = c(
    "year", "overnight service type", "service user count",
    "occupancy rate beds", "occupancy rate rooms"
  ),
  align = "ccc",
  caption = "Table 1: Sample of the key variables in Occupancy & Capacity Data"
)
```

@@

#  Create a subset includes key variables
summary_stats <- data.frame(
  Variable = c(
    "Service User Count", "Occupancy Rate of Beds",
    "Occupancy Rate of Rooms"
  ),
  Mean = c(mean_user_cnt, mean_rate_bed, mean_rate_room),
  SD = c(sd_user_cnt, sd_rate_bed, sd_rate_room)
)

# Present the sample as a table
kable(summary_stats,
  col.names = c("Variable", "Mean", "Standard Deviation"),
  align = "ccc",
  caption = "Table 2: Mean and Standard Deviation of Key Variables"
)
```

@@

#  Create a subset includes key variables
summary_stats <- data.frame(
  Year = c("2021", "2022", "2023", "2024"),
  Service_User_Counts_Sum = c(
    sum_cnt_2021, sum_cnt_2022,
    sum_cnt_2023, sum_cnt_2024
  ),
  Actual_Bed_Capacity_Sum = c(
    sum_bed_2021, sum_bed_2022,
    sum_bed_2023, sum_bed_2024
  ),
  Actual_Room_Capacity_Sum = c(
    sum_room_2021, sum_room_2022,
    sum_room_2023, sum_room_2024
  )
)

# Present the sample as a table
kable(summary_stats,
  col.names = c(
    "Year", "Total Service User Counts",
    "Total Actual Bed Capacity", "Total Actual Room Capacity"
  ),
  align = "ccc",
  caption = "Table 3: Total Service User Count and Actual Capacity between years"
)
```

@@

data |> ggplot(aes(x = overnight_service_type, y = service_user_count)) +
  geom_col(fill = "orange") +
  theme_minimal() +
  labs(x = "overnight service types", y = "service user counts") +
  coord_flip() +
  scale_y_continuous(labels = scales::comma)
```

@@

# Reshape the data to long format for plotting with side-by-side bars
df_long <- tidyr::pivot_longer(avg_rate,
  cols = c("avg_rate_bed", "avg_rate_room"),
  names_to = "type",
  values_to = "average"
)

# Plot using ggplot2
ggplot(df_long, aes(x = factor(year), y = average, fill = type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(x = "Year", y = "Average Occupancy Rate", fill = "Type") +
  theme_minimal() +
  geom_text(aes(label = round(average, 1)),
    position = position_dodge(width = 0.9),
    vjust = -0.5, size = 3.5
  )
```

@@

# Reshape the data to long format for plotting with side-by-side bars
df_long <- tidyr::pivot_longer(total_cap,
  cols = c("total_cap_bed", "total_cap_room"),
  names_to = "type",
  values_to = "total"
)

# Plot using ggplot2
ggplot(df_long, aes(x = factor(year), y = total, fill = type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(x = "Year", y = "Total Actual Capacity", fill = "Type") +
  theme_minimal()
```
